First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-08-04 referee_response.txt

This effect is distinct from other wedge features discussed in the literature, and we have added a paragraph (3rd paragraph of Sec 4.2) pointing this out. The main difference is that this effect is due to the computational limits of the analysis itself, whereas other effects are inherent in the data. In the specific case of Hazelton et al. (2013), the effect is due to multiple baselines migrating in and out of uv cells, and the effective weighting of a given cell is frequency dependent. While the exact mixing would be dependent on the gridding kernel used, it would still exist in the limit of infinitely precise models because the frequency dependent weighting is due to the inherent chromaticity of the instrument.

======================================================================

---2nd paragraph of Section 4.3. "All other sources were given a spectral index of -0.8". What motivated this choice?

This is the median spectral index of sources based on studies below 1.4 GHz. We now mention this and cite the appropriate references.

@@

======================================================================

---Penultimate paragraph of Section 4.3. The authors find that using the two different foreground models gave different flux scales. Is there a sense in which the flux scale has converged?

This is a bit subtle, but the flux scale difference is in the calibrated data, not in the model itself. The difference is quite small (part in 10^3 in mK^2 units), but gets amplified when we show straight difference plots like we use for the diffuse subtraction demonstration. KGS did not find a significant difference in flux scale of the catalog itself, but using a more complete sky model requires lower amplitude (and more accurate) gains to match the visibilities to the model. We have clarified this point in the text.

@@

======================================================================

---Figures 9 and 10. Figure 10 seems to indicate that there is a "+5" pointing. Why is it not included in Figure 9?

This observing campaign did not actually contain data from a +5 pointing, because that is the point that the EoR1 field has risen higher than EoR0 so we switch fields. We have removed the +5 pointing from the cartoon and explain in the caption.

@@

======================================================================

---Figure 11, bottom panel. Since the "+3" portion seems so tightly clustered and is only excluded from the analysis due to some consistent offset, might this be fixable by adjusting calibration parameters?

While it may be possible to recover more pointings in future analysis, it is likely to require both improved calibration and foreground/instrument modeling. The window power statistic is robust against calibration, and the excess power we see in the N-S polarization at this pointing is likely due to bright large scale structure in our sidelobes (which are different in the two polarizations). We would need to properly model and remove this power in order to use this pointing. The potential to recover pointings is mentioned in Sec. 5.1 (end of 2nd paragraph), and discussed more thoroughly in Sec. 6.1.

@@

======================================================================

---8th paragraph of Section 5.2. I think the vertical lines caused by incomplete uv coverage needs more explanation/discussion. A thinning uv coverage alone should not result in the vertical streaks. It needs to be coupled with increased leakage from within the wedge. The fact that reduced uv coverage causes vertical streaks is therefore also a result of how the power spectrum is estimated (since that + the instrument is what determines the leakage). As an example, one could go with a "scorched earth" style analysis, sacrificing sensitivity to preserve the EoR window by doing single-baseline delay spectrum analyses for the longer baselines. The foregrounds would be contained within the wedge under such a procedure, independent of uv-coverage (since we're already at the extreme of analyzing one baseline at a time). The form of one's power spectrum estimator must therefore play a role here.

This is a good point, but we feel the explanation is better suited for the pipeline section. In the penultimate paragraph of Section 3.3 we now go into more detail of the streaks seen in Fig 2.

@@

======================================================================

---Along the same lines as the previous point, might an accounting of covariances between different power spectrum cells also help here, perhaps allowing for decreased leakage?

Indeed, and this is under development in the eppsilon pipeline. We now mention this at the same place as the previous point.

@@

======================================================================

---5th from last paragraph of Section 5.2. Why an increase in slope of 14% of the horizon line? I'm curious as to the thinking here, particularly versus the alternative I've seen in the literature where a constant delay buffer is added to the horizon line.

The cut in kpar and the increased wedge slope effectively achieve the same thing as the constant delay buffer. Our wedge slope is actually a bit more aggressive than one would get from the delay buffer because we saw leakage consistently above the “horizon” (previous 3 hour integrations were on the cusp of reaching the noise level to see this).

@@

This is simply due to the frequency averaging FHD+eppsilon does at the gridding step, which RTS+CHIPS does not do. This is now explained in the caption of Fig. 15.

======================================================================

---It would be helpful in the discussion section to give a sense for what data and results we can look forward to. As an outsider, for example, I am curious as to whether there is more data "in the can" that is just waiting to be analyzed, or whether any new results will come from data taken once the MWA expansion is past its construction/commissioning phase.

We have injected a paragraph (2nd to last of the paper) providing this information. There is plenty of data available, but real progress will be made through understanding the systematics discussed in the paper.

### 2016-08-04 ms.tex
Addressed the remainder of major referee comments. Just one more minor comment to go.

An example two dimensional residual power spectrum formed from the three hour golden 
data set is shown in Figure~\ref{fig:example_ps}. The bulk of the residual power is in the 
so-called foreground ``wedge", indicated with the diagonal black lines where the solid line 
corresponds to sky emission at the horizon, and the dashed lines corresponds to emission 
at the edge of the MWA field of view. Above the wedge we see horizontal lines of 
contamination due to the periodic frequency sampling function. These lines are often 
referred to as the coarse band harmonics. We see vertical streaks at high $k_{||}$ and high 
$k_{\perp}$. These are due to sparse $(u,v)$ sampling beyond the dense core of the MWA 
(starting $\sim70\lambda$), and the foreground power mixing to high spectral 
modes \citep{Bowman:2009}. 
\textbf{We provide axes more closely related to the measurements to assist in connecting
these instrumental effects. The right axis shows delay (Fourier dual to frequency), and the
top axis shows baseline length measured in wavelengths.}
Between the coarse band lines and to the left of the vertical streaks we see regions 
which contain both positive and negative (indicated by blue on the color bar) pixels. These 
regions are where we have successfully isolated the foregrounds and retained a noise-like 
EoR window in the three hour integration.

@@

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][\textbf{N.}~Hurley-Walker et al., in review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source flux densities are prioritized in this order based on our 
confidence in their predicted flux density at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux density from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$. The spectral index was used to extrapolate the flux density from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

@@

Because we use the input foreground model to calibrate our visibilities, the gain estimates 
for our two runs differed. In particular we saw the overall flux scale of the calibrated 
visibilities was higher when using the full hierarchical catalog. In order to put both residual 
2D power spectra onto the same scale for comparison, we first divide each 
\textbf{$(k_{\perp},k_{||})$ cell}
by the 
corresponding dirty power spectrum pixel, then subtract to arrive at a ratio difference. 
\begin{equation}\label{eq:diff_ratio}
\mathtt{ratio\;difference = \frac{residual_1}{dirty_1} - \frac{residual_2}{dirty_2}}
\end{equation}
Here we use subscript 1 to represent the MWACS catalog spectra, and 2 for the 
hierarchical catalog spectra. The ratio difference is shown in Figure~\ref{fig:mwacs_ratio}. 
The entire wedge being positive (blue) demonstrates that using the new catalog resulted in 
successfully subtracting higher fractional power. Not only does this indicate we subtract 
more power, but it also confirms that our calibrated data is more closely matched to the 
model, meaning that our calibration solutions in general are more accurate.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{pointing_cartoon.pdf}
\caption{This is a cartoon depiction of the numbering scheme used to label pointings in
Figures~\ref{fig:jackknife} and \ref{fig:wedge_cut}. Looking South, a field transits from the
East (left), and the telescope begins observing about 5 pointings before zenith (-5). As the
field drifts overhead, we periodically repoint the telescope to recenter the field. Each shift
increments the pointing label by one, and the zenith pointing is defined as zero.
\label{fig:pointing}
}
\end{center}
\end{figure}

@@

>>>>

An example two dimensional residual power spectrum formed from the three hour golden 
data set is shown in Figure~\ref{fig:example_ps}. The bulk of the residual power is in the 
so-called foreground ``wedge", indicated with the diagonal black lines where the solid line 
corresponds to sky emission at the horizon, and the dashed lines corresponds to emission 
at the edge of the MWA field of view. Above the wedge we see horizontal lines of 
contamination due to the periodic frequency sampling function. These lines are often 
referred to as the coarse band harmonics. We see vertical streaks at high $k_{||}$ and high 
$k_{\perp}$. These are due to sparse $(u,v)$ sampling beyond the dense core of the MWA 
(starting $\sim70\lambda$), 
\textbf{which results in non-uniform spectral sampling after gridding. This in turn
causes foreground power to mix to high spectral modes} \citep{Bowman:2009}. 
\textbf{The frequency dependent sampling creates a point spread function in the $k_{||}$
for each $(u,v)$ cell. In principle covariance weighting along the frequency dimension
may be able to mitigate this leakage \citep[e.g.][]{Liu:2011}, and an \eppsilon implementation 
is under development.}
\textbf{We provide axes more closely related to the measurements to assist in connecting
these instrumental effects. The right axis shows delay (Fourier dual to frequency), and the
top axis shows baseline length measured in wavelengths.}
Between the coarse band lines and to the left of the vertical streaks we see regions 
which contain both positive and negative (indicated by blue on the color bar) pixels. These 
regions are where we have successfully isolated the foregrounds and retained a noise-like 
EoR window in the three hour integration.

@@

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][\textbf{N.}~Hurley-Walker et al., in review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source flux densities are prioritized in this order based on our 
confidence in their predicted flux density at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux density from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$, \textbf{the previously determined median spectral index of sources below 1.4 GHz
\citep{Oort:1988, Hunstead:1991, DeBreuck:2000, Mauch:2003}.}
The spectral index was used to extrapolate the flux density from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

@@

Because we use the input foreground model to calibrate our visibilities, the gain estimates 
for our two runs differed. In particular we saw the overall flux scale of the calibrated 
visibilities was higher when using the full hierarchical catalog,
\textbf{owing to a more complete sky model requiring lower amplitude gains to describe the data.}
In order to put both residual 
2D power spectra onto the same scale for comparison, we first divide each 
\textbf{$(k_{\perp},k_{||})$ cell}
by the 
corresponding dirty power spectrum pixel, then subtract to arrive at a ratio difference. 
\begin{equation}\label{eq:diff_ratio}
\mathtt{ratio\;difference = \frac{residual_1}{dirty_1} - \frac{residual_2}{dirty_2}}
\end{equation}
Here we use subscript 1 to represent the MWACS catalog spectra, and 2 for the 
hierarchical catalog spectra. The ratio difference is shown in Figure~\ref{fig:mwacs_ratio}. 
The entire wedge being positive (blue) demonstrates that using the new catalog resulted in 
successfully subtracting higher fractional power. Not only does this indicate we subtract 
more power, but it also confirms that our calibrated data is more closely matched to the 
model, meaning that our calibration solutions in general are more accurate.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{pointing_cartoon.pdf}
\caption{This is a cartoon depiction of the numbering scheme used to label pointings in
Figures~\ref{fig:jackknife} and \ref{fig:wedge_cut}. Looking South, a field transits from the
East (left), and the telescope begins observing about 5 pointings before zenith (-5). As the
field drifts overhead, we periodically repoint the telescope to recenter the field. Each shift
increments the pointing label by one, and the zenith pointing is defined as zero.
\textbf{While a +5 pointing (and beyond) is possible with the MWA, this observing campaign
did not contain any such data because we instead switched to the EoR1 field at that time.}
\label{fig:pointing}
}
\end{center}
\end{figure}

@@

\textbf{Between 2013 and the time of writing, the MWA has collected over two thousand 
hours of data targeting the EoR fields. These observations include the three designated sky 
fields and two frequency bands. In addition, observations are underway to further 
characterize the sources in the instrument sidelobes. The second phase of the MWA
will be commissioned in the latter half of 2016, and will include two sets of highly redundant
cores of antennas. This hybrid configuration will enable a unique opportunity to study
the calibration techniques used in imaging and delay spectrum analyses, furthering
our understanding of the systematics limiting our current results.}
